Batch, Off-Policy Apprencticeship Learning

### 2011-03-28 main.tex
Intro de Matthieu

\title{Batch, Off-Policy Apprencticeship Learning}
\author{Edouard Klein \and Matthieu Geist \and Olivier Pietquin \\
Supélec, IMS\\
Metz, France \\
\{edouard.klein$|$matthieu.geist$|$olivier.pietquin\}@supelec.fr}
\begin{document}
\maketitle
\begin{abstract}
We consider the IRL framework for apprenticeship learning. %1
Current IRL algorithms use a simple Monte-Carlo estimation to approximate the feature expectation of both the expert and the policy under evaluation. %2
In this paper, we propose another method for estimating the feature expectation based on the Least Square Temporal Difference (LSTD) family of algorithms. %3
This new method allows for batch, off-policy appreticeship learning in situations where the dynamics of the system remains unknown. %4
By testing it on a classical problem, we show that our method is empirically sound.%5
We also provide some theoretical results.%6
\end{abstract}
\section{Introduction}
The optimal control problem can be expressed as having a machine control a system, with the goal of fulfilling a specific task. Various solutions to this problem exist. Things can get hard to deal with, however, when precisely defining the task is in itself a challenge. When the system can be empirically controlled by an expert, even when his behaviour is difficult to describe formally, apprenticeship learning is a way to let a machine control the system by mimicking the expert.\\

When the task at hand can be formally described, the problem of optimal control can be solved by tools from the \emph{Reinforcement Learning} (RL) framework. An extensive introduction to this matter can be found in \citep{sutton1998reinforcement} : the control problem is projected into a sequential decision problem inside a \emph{Markov Decision Process}. The agent transits from \emph{state} to \emph{state}, the decision taking the form of choosing an \emph{action} at each visited state. For each \emph{transition}, the agent receives a \emph{reward}. The \emph{reward function} (a mapping from each state to a real value) defines the task. The agent's goal is to maximize the long-term reward by maximizing what is called the \emph{expected discounted reward} : the expectation of the sum of all rewards to be obtained, multiplied by a deacreasing-over-time factor (called the \emph{discount factor}). For a given \emph{policy} (a mapping from each state to an action) that the agent follows, the mapping from each state to the expected discounted reward is called the \emph{value function}. The \emph{optimal policy}, hopefully found by the agent, is the unique policy of which the value fuction is, for every state, superior or equal to all the other policie's value functions.\\

It is not necessary for a model of the environment or for a simulator to be available, as for example the LSTD (\citep{bradtke1996linear}) algorithm can approximate the value function in a batch, off-policy, model-free setting. Building on it, the LSPI (\citep{lagoudakis2003least}) algorithm can be used to actually solve the RL problem in such settings : given a set of random or arbitrary transitions and the corresponding rewards, find the optimal policy.\\

\citet{russell1998learning} first defined the \emph{Inverse Reinforcement Learning} (IRL) problem in the terms used today, drawing from animal behavior research and more loosely from macroeconomics research. What is looked for is not the policy as it is in the RL problem, but the reward, which defines the task and of which the expected discounted sum is maximized by an expert. IRL is a framework in which it could be possible to attack the apprenticeship learning problem mentionned earlier.\\

\citet{ng2000algorithms} first proposed an algorithm to solve the IRL problem, sheding light on a problem foreseen in \citep{russell1998learning} : the IRL problem is mathematically ill-defined as it admits a infinity of solutions, including degenerative ones (e.g. the null reward accepts any policy as an optimal policy).\\

This work was followed by an article \citep{abbeel2004apprenticeship} focussing on the more specific, but better defined problem of apprenticeship learning using the IRL framework. This work makes use of the IRL framework by using an approximator for the reward function, and looking for the parameters for which the approximated reward function is the best. This however only concerns the apprenticeship learning (a.k.a. imitation learning) because the metric used, although intended to discriminate the expert's performance and the current policy's performance \emph{under the current approximate reward}, can be seen as a measure of proximity of the different states visitation frequencies.

This metric, called the \emph{feature expectation}, is akin to the expected discounted reward discussed when describing the RL framework. It does however not sum the reward but the features of the visited states, features defined by the choice of the reward function approximator. It can be assessed either by analytical resolution (when the model is available) of a Monte-Carlo estimation (when a simulator is available).\\

The main limitation in \citet{abbeel2004apprenticeship}'s work (from now on reffered to as the ANIRL algorithm) is the need for a model of the environnement or, at least, a simulator. We build directly upon this work by presenting a new algorithm, LSTD$\mu$, which along with the use of LSPI and LSTD allows the use of the ANIRL algorithm in a batch, off-policy, model-free setting.\\

FIXME:Autres approches qu'Abbeel, notamment celles qui cherchent à contourner le même problème.\\

@@

The main shortcoming of this algorithm lies in its need for either a model of the environnement or at least a sample source $D$, in order to approximate the feature expectation of both the expert and the current policy via a Monte-Carlo estimator.
\section{LSTD-$\mu$}
The similarity between equations \ref{eqn:mu} and \ref{eqn:V} (both are Bellman equations) allows the use of the LSTD algorithm in order to estimate the feature expectation of a policy.\\

@@

\section{Analysis}
FIXME: Matthieu ?
\section{Experimental benchmark}
The experimental benchmark chosen here is the same as in \citep{ng2000algorithms}, a 5x5 grid world. The agent is in one of the case of the grid (whose coordinates is the state) and can choose at each step one of the four compass directions (the action). With probability 0.9, the agent moves in the intended direction. With probability 0.1, the direction is randomly drawn (and thus have probability 0.25 to match the original direction). The reward optimized by the expert is 0 everywhere but in the upper-right square, where it is 1. For every policy, an episodes ends when the upper right square is attained, or after 20 steps. At the start of each episode, the agent is in the lower-left corner of the grid (the opposite of where the reward is).\\

@@

It shall be noted that although it plays its role, the halt criterion is not a good measure of the quality of the current policy in mc\_ANIRL either, as it can be low when the policy is bad. The best policy, however, can be easily chosen from the $||\mu(s_0) - \mu_E(s_0)||_2$ term, which mc\_ANIRL can compute. When this term is low, the objective performance (that is, $V^E(s_0)-V^\pi(s_0)$ with respect to the unknow true reward function) is low too. 
\begin{figure}
\includegraphics[width=.5\textwidth]{../Code/courbe_e.pdf}
\caption{$||\mu_E(s_0)-\mu_\pi(s_0)||_2$ with respect to the number of samples available from the expert. Our algorithm is not as good as \citet{abbeel2004apprenticeship} with very few samples, but make up for it very soon.}
\label{fig:E}
\end{figure}
\label{sec:perf}
\begin{figure}
\includegraphics[width=.5\textwidth]{../Code/courbe_a.pdf}
\caption{Different criteria with respect to the number of iterations for a run of mc\_ANIRL.}
\label{fig:A}
\end{figure}
\begin{figure}
\includegraphics[width=.5\textwidth]{../Code/courbe_b.pdf}
\caption{Different criteria with respect to the number of samples from the expert, for several runs of lstd\_ANIRL.}
\label{fig:B}
\end{figure}
\section{Future Work}
FIXME: Une seule itération, est-ce normal ? Abbeel reste-t-il meilleur sur un problème plus complexe ? Transfert de tâche ? Est-il nécessaire de choisir sa politique avec LSTDMu, puisqu'il finit en une seule fois ?
\section{Conclusion}
\bibliographystyle{named}
\bibliography{../Biblio/Biblio}
\end{document}

>>>>

\title{Batch, Off-Policy Apprencticeship Learning}
\author{Edouard Klein \and Matthieu Geist \and Olivier Pietquin \\
Supélec, IMS\\
Metz, France \\
\{edouard.klein$|$matthieu.geist$|$olivier.pietquin\}@supelec.fr}
\begin{document}
\maketitle
\begin{abstract}
We consider the IRL framework for apprenticeship learning. %1
Current IRL algorithms use a simple Monte-Carlo estimation to approximate the feature expectation of both the expert and the policy under evaluation. %2
In this paper, we propose another method for estimating the feature expectation based on the Least Square Temporal Difference (LSTD) family of algorithms. %3
This new method allows for batch, off-policy appreticeship learning in situations where the dynamics of the system remains unknown. %4
By testing it on a classical problem, we show that our method is empirically sound.%5
We also provide some theoretical results.%6
\end{abstract}
\section{Introduction}
The optimal control problem can be expressed as having a machine
control a system, with the goal of fulfilling a specific task.
Various solutions to this problem exist. Notably, the reinforcement
learning (RL) paradigm~\citep{sutton1998reinforcement} is a general framework in
which an agent learns to control optimally a dynamic system through
interactions with it. The task is specified through a reward
function, the agent objective being to take sequential decisions so
as to maximize the expected cumulative reward.

However, things can get hard to deal with  when precisely defining
the task is in itself a challenge. When the system can be
empirically controlled by an expert, even when its behavior is
difficult to describe formally, apprenticeship learning is a way to
let a machine control the system by mimicking the expert. Rather
than directly mimicking the expert with some supervised learning
approach, inverse reinforcement learning (IRL)~\citep{ng2000algorithms}
consists in learning a reward function under which the policy
demonstrated by an expert is optimal. Mimicking the expert therefore
ends up to learning an optimal policy according to this reward
function. A significant advantage of such an approach is that
expert's actions can be guessed in states which have not been
encountered during demonstration.

Basically, there are three families of IRL algorithms:
feature-expectation-based~\citep{abbeel2004apprenticeship,syed2008game,syed2008apprenticeship,ziebart2008maximum,boularias2011bootstrapping},
marge-maximization-based~\citep{ratliff2006maximum,ratliff2007boosting,ratliff2007imitation,kolter2008hierarchical}
and approaches based on the parameterization of the policy by the
reward function~\citep{Ramachandran2007bayesian,Neu2007apprenticeship}. The first family
assumes a linearly parameterized reward function. This naturally
leads to a linearly parameterized value function, the associated
feature vector being the so-called feature expectation (defined as
the cumulative discounted reward's feature vector under a given
policy, see Section~\ref{sec:back} for a formal definition).
These approaches learn a reward function such that the feature
expectation of the optimal policy (according to the learnt reward
function) is close to the feature expectation of the expert policy.
This is a sufficient condition to have close value functions. The
second family expresses IRL as a constrained optimization problem in
which expert's examples have higher expected reward than all other
policies by a certain margin. Moreover, suboptimality of the expert
can be considered through the introduction of slack variables. The
last family parameterizes policies with a reward function. Assuming
that the expert acts according to a Gibbs policy (respectively to
the optimal value function related to the reward function which is
optimized), it is possible to estimate the likelihood of a set of
state-action pairs provided by the expert. The algorithms differs in
the way this likelihood is maximized.

This paper focuses on the first family of algorithms, and more
precisely on its founding algorithm~\citep{abbeel2004apprenticeship}. All of them
rely on the computation of the feature expectation (which depends on
policies but not on rewards) of (i) the expert and (ii) some
intermediate policies. The expert's feature expectation is computed
using a simple Monte Carlo approach (which requires full
trajectories of the expert). Other feature expectations are either
computed exactly (which requires knowing analytically the dynamic of
the system) or with a Monte Carlo approach (which requires to
simulate the system). The contribution of this paper is LSTD-$\mu$,
a new temporal-difference-based algorithm to estimate these feature
expectations. It relaxes the predeceasing assumptions: transitions
of the expert are sufficient (rather than full trajectories) and nor
the model neither a simulator are necessary to compute intermediate
feature expectations. This paper focuses on the algorithm introduced
in~\citep{abbeel2004apprenticeship}, but we are confident that the proposed
approach can be extended to other algorithms based on feature
expectation computation~\citep{syed2008game,syed2008apprenticeship,ziebart2008maximum}.

The rest of this paper is organized as follows.
Section~\ref{sec:back} provides the necessary background,
notably the definition of feature expectation and its use in the
seminal IRL algorithm of~\citep{abbeel2004apprenticeship}.
Section~\ref{sec:lstdmu} presents LSTD-$\mu$, our main contribution.
Section~\ref{sec:exp} provides some preliminary experiments and
Section~\ref{sec:conclusion} opens perspectives.

@@

The main shortcoming of this algorithm lies in its need for either a model of the environnement or at least a sample source $D$, in order to approximate the feature expectation of both the expert and the current policy via a Monte-Carlo estimator.
\section{LSTD-$\mu$}
\label{sec:lstdmu}
The similarity between equations \ref{eqn:mu} and \ref{eqn:V} (both are Bellman equations) allows the use of the LSTD algorithm in order to estimate the feature expectation of a policy.\\

@@

\section{Analysis}
FIXME: Matthieu ?
\section{Experimental benchmark}
\label{sec:exp}
The experimental benchmark chosen here is the same as in \citep{ng2000algorithms}, a 5x5 grid world. The agent is in one of the case of the grid (whose coordinates is the state) and can choose at each step one of the four compass directions (the action). With probability 0.9, the agent moves in the intended direction. With probability 0.1, the direction is randomly drawn (and thus have probability 0.25 to match the original direction). The reward optimized by the expert is 0 everywhere but in the upper-right square, where it is 1. For every policy, an episodes ends when the upper right square is attained, or after 20 steps. At the start of each episode, the agent is in the lower-left corner of the grid (the opposite of where the reward is).\\

@@

It shall be noted that although it plays its role, the halt criterion is not a good measure of the quality of the current policy in mc\_ANIRL either, as it can be low when the policy is bad. The best policy, however, can be easily chosen from the $||\mu(s_0) - \mu_E(s_0)||_2$ term, which mc\_ANIRL can compute. When this term is low, the objective performance (that is, $V^E(s_0)-V^\pi(s_0)$ with respect to the unknow true reward function) is low too. 
\begin{figure}
\includegraphics[width=.5\textwidth]{../Code/courbe_e.pdf}
\caption{$||\mu_E(s_0)-\mu_\pi(s_0)||_2$ with respect to the number of samples available from the expert. Our algorithm is not as good as \citet{abbeel2004apprenticeship} with very few samples, but make up for it very soon.}
\label{fig:E}
\end{figure}
\label{sec:perf}
\begin{figure}
\includegraphics[width=.5\textwidth]{../Code/courbe_a.pdf}
\caption{Different criteria with respect to the number of iterations for a run of mc\_ANIRL.}
\label{fig:A}
\end{figure}
\begin{figure}
\includegraphics[width=.5\textwidth]{../Code/courbe_b.pdf}
\caption{Different criteria with respect to the number of samples from the expert, for several runs of lstd\_ANIRL.}
\label{fig:B}
\end{figure}
\section{Conclusion}
\label{sec:conclusion}
\bibliographystyle{named}
\bibliography{../Biblio/Biblio}
\end{document}
